---
title: "Assignment 1 - Language Development in ASD - part 4"
author: "Riccardo Fusaroli"
date: "August 10, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Welcome to the fourth exciting part of the Language Development in ASD exercise

In this exercise we will assess how many participants we would need to adequately replicate our findings (ensuring our sample size is adequate, our alpha at 0.05 and our beta at 0.8).

### Exercise 1

How much power does your study have (if your model estimates are quite right)?
- [GitHub]Load your dataset, fit your favorite model, assess power for your main effects and interactions of interest.
- Report the power analysis and comment on what you can (or cannot) use its estimates for.
```{r}
#setwd
setwd("C:/Users/Samsung/Desktop/expmeth/Assignment-4")

#libraries
library(ggplot2)
library(MASS)
library(dplyr)
library(lme4)
library(MuMIn)
library(lmerTest)
library(gstat)
library(stringr)
library(plyr)
library(caret)
library(modelr)
library(ModelMetrics)
library(Metrics)
library(tidyverse)
library(simr)
library(pacman)

#Round to four digits
options(digits = 4)

#load dataset
traind <- read.csv("cleandata.csv")
testd <- read.csv("testdata.csv")
testd <- subset(testd, select = -c(X)) #dataset without x
traind<- subset(traind, select= -c(X))

#Change Subject ID's to not overlap
testd$SUBJ = testd$SUBJ+61

#merge datasets
data = rbind(traind, testd)

#factors
data$SUBJ=as.factor(data$SUBJ)
data$Diagnosis=as.factor(data$Diagnosis)

#Recreate favourite model
m <- lmer(CHI_MLU~VISIT+Diagnosis+types_CHI+verbalIQ+tokens_CHI+(1+VISIT|SUBJ), data = data) 

#fit the model
summary(m) #output
r.squaredGLMM(m) #r2m=0.7375

#Test power with current beta values for each variable
powerVISIT = powerSim(m, fixed("VISIT"), nsim=100)
powerDiagnosis = powerSim(m, fixed("DiagnosisTD"), nsim=100)
powertypes_CHI = powerSim(m, fixed("types_CHI"), nsim=100)
powerverbalIQ = powerSim(m, fixed("verbalIQ"), nsim=100)
powertokens_CHI = powerSim(m, fixed("tokens_CHI"), nsim=100)

powerVISIT #98%, effect size 0.080, alpha=0.05, nrow=387
powerDiagnosis #0%, effect size 0.080,  alpha=0.05, nrow=387
powertypes_CHI #100%, effect size 0.0070, alpha=0.05, nrow=387
powerverbalIQ #93%, effect size 0.023, alpha=0.05, nrow=387
powertokens_CHI #19%, effect size 0.00022,  alpha=0.05, nrow=387

#how to report it? what can i use it for?

```


### Exercise 2

How would you perform a more conservative power analysis?
- Identify and justify a minimum effect size for each of your relevant effects
- [GitHub] take the model from exercise 1 and replace the effects with the minimum effect size that you'd accept.
- [GitHub] assess the power curve by Child.ID, identifying an ideal number of participants to estimate each effect
- OPTIONAL if your power estimates do not reach an acceptable threshold simulate additional participants and repeat the previous analysis
- Report the power analysis and comment on what you can (or cannot) use its estimates for.

```{r}
# Change to lowest relevant effect sizes.. (don't know how to choose now, leave effect sizes approx. as they are)
model<-m
fixef(model)
fixef(model)["VISIT"] <- 0.05
fixef(model)["DiagnosisTD"] <- 0.07
fixef(model)["types_CHI"] <- 0.006
fixef(model)["verbalIQ"] <- 0.02
fixef(model)["tokens_CHI"] <- 0.0002


#Calculate the power depending on number of participants
powerCurveVISIT = powerCurve(model, fixed("VISIT"), along="SUBJ", nsim=50)
powerCurveDiagnosis = powerCurve(model, fixed("DiagnosisTTD"), along="SUBJ", nsim=50)
powerCurvetypes_CHI = powerCurve(model, fixed("types_CHI"), along="SUBJ", nsim=50)
powerCurveverbalIQ = powerCurve(model, fixed("verbalIQ"), along="SUBJ", nsim=50)
powerCurvetokens_CHI = powerCurve(model, fixed("tokens_CHI"), along="SUBJ", nsim=50)


#And Plot it
plot(powerCurveVISIT) #close to 60
plot(powerCurveDiagnosis)#it's gonna be zero 
plot(powerCurvetypes_CHI) #15
plot(powerCurveverbalIQ) #40+, 65 for sure
plot(powerCurvetokens_CHI) #way more than 60


### Riccardo's clumsy function to simulate new participants
### TO DO points are only notes for myself, so not part of the assignment

createNewData <- function (participants,visits,model){
  # participants is the number of subjects
  # visits is the number of visits
  # TO DO: LOOP THROUGH ALL FE ROWS AND AUTOMATICALLY EXTRACT NAMES OF FIXED EFFECTS AND ESTIMATES
  fe <- fixef(model)
  Intercept <- fe[1] #intercept
  bVisit <- fe[2] #visit
  bDiagnosis <- fe[3] #diagnosis
  bVisitDiagnosis <- fe[4] #visit diagnosis interaction
  # TO DO: INTEGRATE STANDARD ERROR?
  
  # TO DO: LOOP THROUGH ALL VC COMPONENTS AND AUTOMATICALLY EXTRACT NAMES OF EFFECTS AND ESTIMATES
  vc<-VarCorr(model) # variance component
  sigmaSubject <- as.numeric(attr(vc[[1]],"stddev")[1]) # random intercept by subject
  sigmaVisit <- as.numeric(attr(vc[[1]],"stddev")[2]) # random slope of visit over subject
  sigmaResiduals <- as.numeric(attr(vc,"sc"))
  sigmaCorrelation <- as.numeric(attr(vc[[1]],"correlation")[2])
  
  # Create an empty dataframe
  d=expand.grid(VISIT=1:visits,SUBJ=1:participants)
  # Randomly sample from a binomial (to generate the diagnosis)
  condition <- sample(rep(0:1, participants/2))
  d$Diagnosis<-condition[d$SUBJ]
  d$Diagnosis[is.na(d$Diagnosis)]<-1
  
  ## Define variance covariance matrices:
  Sigma.u<-matrix(c(sigmaSubject^2,
                    sigmaCorrelation*sigmaSubject*sigmaVisit,
                    sigmaCorrelation*sigmaSubject*sigmaVisit,
                    sigmaVisit^2),nrow=2)
  
  ## generate new fake participants (column1=RandomIntercept, column2=RandomSlope)
  u<-mvrnorm(n=participants,
             mu=c(0,0),Sigma=cov(ranef(model)$SUBJ))
  
  ## now generate fake data:
  ### the outcome is extracted from a gaussian with
  ### the solution to the model's equation as mean and
  ### the residual standard deviation as standard deviation 
  d$CHI_MLU <- rnorm(participants*visits,
                     (Intercept+u[,1]) +
                     (bVisit+u[,2])*d$VISIT + 
                     bDiagnosis*d$Diagnosis ,sigmaResiduals)  
  
  return(d)
}


```


### Exercise 3

Assume you have only the resources to collect 30 kids (15 with ASD and 15 TDs). Identify the power for each relevant effect and discuss whether it's worth to run the study and why.

```{r}
#Simulate dataset
d_sim = createNewData(500, 6, m)
d_sim$SUBJ=as.factor(d_sim$SUBJ)

#Train model on simulated data
Model_Sim = lmer(CHI_MLU ~ VISIT*Diagnosis + (1+VISIT|SUBJ), d_sim)
summary(Model_Sim)


#Run a basic model to find smallest interesting effect for the model on the simulated data
Model_0 = lmer(CHI_MLU ~ VISIT*Diagnosis + (1+VISIT|SUBJ), data)
summary(Model_0)
fixef(Model_0)


#Define smallest interesting effect size
fixef(Model_Sim)["VISIT"] <- 0.1
fixef(Model_Sim)["Diagnosis"] <- -0.2
fixef(Model_Sim)["VISIT:Diagnosis"] <- 0.2


#Calculate power for different amounts of participants
powerCurveVISITsim = powerCurve(Model_Sim, fixed("VISIT"), along="SUBJ", nsim=50)
powerCurveDiagnosissim = powerCurve(Model_Sim, fixed("Diagnosis"), along="SUBJ", nsim=50)
powerCurveinteractsim = powerCurve(Model_Sim, fixed("VISIT:Diagnosis"), along="SUBJ", nsim=50)

plot(powerCurveVISITsim)
plot(powerCurveDiagnosissim)
plot(powerCurveinteractsim)
```






